---
title: "Gathering Data"
author: "Ciara Duggan"
date: "10/15/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
 
# Setting up file and loading libraries
library(tidyverse)
library(readr)
```


```{r rx_data}

# Loading AmfAR county-level opioid Rx rate dataset (2006-2017).

county_rx_rate <- 
  read_csv(file = "/Users/ciaraduggan/Desktop/gov_50/final_project/duggan_final_project_gov50/raw_data/county_opioid_RxRate.csv", 
           col_types = cols(
             STATE = col_character(),
             STATEABBREVIATION = col_character(),
             COUNTY = col_character(),
             STATEFP = col_character(),
             COUNTYFP = col_character(),
             INDICATOR = col_character(),
             YEAR = col_character(),
             VALUE = col_double()
))


```

# AmfAR County-Level Opioid Prescription Data (per 100 persons)

This dataset was downloaded from the AmfAR Opioid and Health Indicators database.
The specific page in the database can be found here: 
https://opioid.amfar.org/indicator/opioid_RxRate

For indicator "opioid_RxRate", the "value" measure signifies the number of 
opioid prescriptions per 100 persons in the county.

From AmfAR: "This measure uses IQVIA data from pharmacies which dispense nearly 
90% of all retail prescriptions in the US. It includes prescriptions paid for by
commercial insurance, Medicaid, Medicare, or cash or equivalent. Cough and cold 
formulations containing opioids and buprenorphine products typically used to
treat opioid use disorder are excluded, and IQVIA data does not include 
methadone dispensed through methadone maintenance treatment programs."

Data Source: Centers for Disease Control and Prevention

Available at: https://www.cdc.gov/drugoverdose/maps/rxrate-maps.html

```{r drug_death_data}

# Loading AmfAR data on county-level age-adjusted drug-related death rate (per
# 100,000).

county_drug_deaths <-
  read_csv(file = "/Users/ciaraduggan/Desktop/gov_50/final_project/duggan_final_project_gov50/raw_data/county_ageadjusted_drugdeathrate.csv", 
           col_types = cols(
             STATE = col_character(),
             STATEABBREVIATION = col_character(),
             COUNTY = col_character(),
             STATEFP = col_character(),
             COUNTYFP = col_character(),
             INDICATOR = col_character(),
             YEAR = col_character(),
             VALUE = col_double()
)) 
       
```
  
          
# AmfAR Data on Age-Adjusted Drug Poisoning Deaths per 100,000 (Modeled)

This dataset was downloaded from the AmfAR Opioid and Health Indicators 
database. The specific page in the database can be found here: 
https://opioid.amfar.org/indicator/drugdeathrate_est

For indicator "drugdeathrate_est", the "value" measure signifies the estimated
rate of death from drug poisoning, including both illicit and 
prescription drugs, per 100,000 U.S. standard population for 2000. 

From AmfAR: "Estimates come from the Centers for Disease Control and Prevention
(CDC) National Center for Health Statistics (NCHS) Data Visualization Gallery. 
It should be noted that these data are estimates, and are drawn from a model
designed to generate stable estimates of death rates where data are sparse due
to small population size. Specific county values should be interpreted with 
caution, and the actual measure of drug deaths per 100,000 should be preferred 
where available. More information on the estimation procedure and county data 
including confidence intervals are available at the link below.

Data Source: Centers for Disease Control's National Center for Health Statistics 
Data Visualization Gallery.

Available at:
https://www.cdc.gov/nchs/data-visualization/drug-poisoning-mortality/


```{r data_wrangle}

# Pivoting the county_drug_deaths dataset to a wider format to create a variable
# containing age-adjusted drug-related death rate estimates.

county_drug_deaths_wider <-
  county_drug_deaths %>%
  pivot_wider(names_from = "INDICATOR", values_from = "VALUE")

# Pivoting the county_rx_rate_wider dataset to a wider format to create a
# variable containing the opioid prescription rates per 100,000 people.

county_rx_rate_wider <-
  county_rx_rate %>%
  pivot_wider(names_from = "INDICATOR", values_from = "VALUE")

# Joining the resulting county_drug_deaths_wider and county_rx_rate_wider data
# sets by county and year.

county_drug_data <- county_drug_deaths_wider %>%
  inner_join(county_rx_rate_wider, by = c("COUNTY", 
                                          "YEAR", 
                                          "STATE", 
                                          "STATEABBREVIATION",
                                          "STATEFP",
                                          "COUNTYFP"
  ))
```

# Test plot

```{r test_plot}

# Creating a test scatterplot for year 2006 with opioid prescription rate per 
# 100,000 on x-axis and age-adjusted drug-related deaths per 100,000 on y-axis.

county_drug_data %>%
  filter(YEAR == 2006) %>%
  ggplot(aes(x = opioid_RxRate, y = drugdeathrate_est)) +
  geom_point(alpha = 0.1) +
  geom_smooth(method = lm)

```

# Code

```{r print-code, ref.label=knitr::all_labels(), echo = TRUE, eval=FALSE}
# Printing code at bottom
```

